---
title: "Scotch QOF Analaysis"
subtitle: "Exploratory analysis"
author: "David Henderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    theme: spacelab
    toc: true
    toc_float: true
    df_print: paged
    highlight: haddock
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 12, fig.height = 9,
                      warning = FALSE, message = FALSE,
                      class.source="bg-success")
```


# Introduction

Initial analysis of qof data

Import and clean TQA data - needs to be made long

Have a look at what is measurable back the way. Import those

Then find SIMD and UR etc for each practice

Then start descriptives

## Software

R packages and required helper functions.

```{r, warning=FALSE, message=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(socviz))

source(here("01_functions/davidhen_helper_funcs_settings.R"))

```


# Load and Clean Data

## Post-2016 TQA data

### Indicators

There is a combined 2016-2019 achievement file for TQA, however it is in a wide format.

```{r}
# Load in the file and tidy variable names
tqa_ind <- read_excel(here("03_data/TQA/IR2019-00546 TQA Achievement2016-2019.xlsx"), 
                  sheet = "Data") %>%
  clean_names()

# Split the year variable to lose the unneeded text
# Drop the percentage variables (recalculated later)
# Ensure there is not leading or trailing white space in character variables
tqa_ind %>% 
  separate(year, c("fin_year", "meh"), sep = " ") %>% 
  select(-meh, -ends_with("_perc")) %>% 
  mutate(across(is.character, .fns = str_trim)) -> tqa_ind

# Drop the denominator variables, reshape the numerator variables and values
tqa_ind %>%
  select(-ends_with("_den")) %>% 
  pivot_longer(ends_with("_num"), names_to = "indicator", 
               values_to = "ind_num") %>% 
  mutate(indicator = str_remove(indicator, "_num")) -> tqa_num

# Repeat for the denominators
tqa_ind %>%
  select(-ends_with("_num")) %>% 
  pivot_longer(ends_with("_den"), names_to = "indicator", 
               values_to = "ind_den") %>% 
  mutate(indicator = str_remove(indicator, "_den")) -> tqa_den

# join the two together, recalculate % arrange for an easier quick look
left_join(tqa_num, tqa_den) %>% 
  mutate(ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  arrange(practice, indicator, fin_year) -> tqa_ind

# remove interim objects
rm(list = c("tqa_num", "tqa_den"))

# Print out the result
tqa_ind
```



```{r}
unique(tqa_ind$indicator)
```

25 indicators in the the most recent TQA data. Will keep all that are consistent going back in time.


### Registers

Two files for registers in the TQA data

```{r}
# Import 2016-2018 data
reg_1618 <- read_excel(here("03_data/TQA/IR2018-01954 TQA 2016-2018 Prevalence.xlsx"),
                       sheet = "Data") %>% 
  clean_names() %>% 
  # Tidy variables up and drop prevalence (recalculate later)
  separate(year, c("fin_year", "meh"), sep = " ") %>% 
  select(-meh, -ends_with("prev")) %>% 
  mutate(across(is.character, .fns = str_trim)) %>% 
  # Reshape register values to long format
  pivot_longer(ends_with("reg"), names_to = "disease", 
               values_to = "registered") %>% 
  mutate(disease = str_remove(disease, "reg")) %>%
  # recalculate prevalence - note as a % - same as original .xlsx
  mutate(prev = registered/population * 100) %>% 
  round_df(dig = 1) -> reg_1618

# Repeat for 2018/19 data
reg_1819 <- read_excel(here("03_data/TQA/IR2019-00546 TQA 2018-2019 Prevalence.xlsx"),
                       sheet = "Data") %>% 
  clean_names() %>% 
  separate(year, c("fin_year", "meh"), sep = " ") %>% 
  select(-meh, -ends_with("prev")) %>% 
  mutate(across(is.character, .fns = str_trim)) %>% 
  pivot_longer(ends_with("reg"), names_to = "disease", 
               values_to = "registered") %>% 
  mutate(disease = str_remove(disease, "reg")) %>% 
  mutate(prev = registered/population * 100) %>% 
  round_df(dig = 1) -> reg_1819

# Join them together
bind_rows(reg_1618, reg_1819) %>% 
  arrange(practice, fin_year) -> tqa_reg

#remove interim objects

rm(list = c("reg_1618", "reg_1819"))

# Print table
tqa_reg
```


## Pre-2016 QOF data‚

### Practice populations

Practice populations are annoyingly kept in a different file from the indicators and registers. Will import and tidy up here so we have a table that can be joined later on.

```{r}
# Import the 2015/16 lookup file for practice list size
pop_1516 <- 
  read_excel(here("03_data/QOF_2015_16/QOF_Scot_201516_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201516") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population),
         fin_year = "2015-16")

pop_1415 <- 
  read_excel(here("03_data/QOF_2014_15/Individual level 2014-5/QOF_Scot_201415_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201415") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population),
         fin_year = "2014-15")

pop_1314 <- 
  read_excel(here("03_data/QOF_2013_14/Individual level 2013-14/QOF_Scot_201314_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201314") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population), 
         fin_year = "2013-14")

pop_1213 <- 
  read_excel(here("03_data/QOF_2012_13/Individual level 2012-13/QOF_Scot_201213_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201213") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population), 
         fin_year = "2012-13")

pop_1112 <- 
  read_excel(here("03_data/QOF_2011_12/Individual level 2011-12/QOF_Scot_201112_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201112") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population), 
         fin_year = "2011-12")

pop_1011 <- 
  read_excel(here("03_data/QOF_2009_10/Individual level 2009-10/QOF_Scot_201011_Practice_lookup.xls"),
             sheet = "Practices_in_QOF_201011") %>% 
  clean_names %>% 
  select(practice = practice_code, population = list_size) %>% 
  mutate(practice = as.character(practice), 
         population = as.double(population),
         fin_year = "2010-11")

pop_table <- bind_rows(pop_1516, pop_1415, pop_1314, pop_1213, pop_1112,
                       pop_1011) %>% 
  arrange(practice, fin_year) %>% 
  select(fin_year, everything())

# remove interim objects
rm(list = c("pop_1516", "pop_1415", "pop_1314", "pop_1213",
            "pop_1112", "pop_1011"))

pop_table
```


### AF

#### Indicator

The AF indicator in TQA is "af007". 

> In those patients with atrial fibrillation with a record of a CHA2DS2-VASc score of 2 or more, the percentage of patients who are currently treated with anti-coagulation drug therapy.

There are slight variations on the name in some years e.g. "AF007(S)" or "AF04" which isn't too big of a problem. However, there is a little discrepancy before 2011/12 where "AF03" included anti-platelet therapy as well as anti-coagulation therapy. Too big of change? Bruce and Dan to answer. 

```{r}
# Import AF file for 2015/16, tidy and rename for consistency with TQA data, 
# Add a variable indicating the financial year
# Join with list size table created above
#af_1516 <-
  read_excel(here("03_data/QOF_2015_16/QOF_Scot_201516_indiv_af.xls"),
                      sheet = "Data Table") %>% 
  clean_names() %>% 
  filter(ind == "AF007(S)") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2015-16", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)

# Repeat for 2014/15 indicator "AF004"
af_1415 <-
  read_excel(here("03_data/QOF_2014_15/Individual level 2014-5/QOF_Scot_201415_indiv_af.xls"),
                      sheet = "Data Table") %>% 
  clean_names() %>% 
  filter(ind == "AF004") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2014-15", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)

# Repeat for 2013/14 indicator "AF004"
af_1314 <-
  read_excel(here("03_data/QOF_2013_14/Individual level 2013-14/QOF_Scot_201314_indiv_af.xls"),
           sheet = "Data Table") %>% 
  clean_names() %>% 
  filter(ind == "AF004") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2013-14", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)

# Repeat for 2012/13 indicator "AF07"
af_1213 <- 
  read_excel(here("03_data/QOF_2012_13/Individual level 2012-13/QOF_Scot_201213_indiv_af.xls"),
           sheet = "Atrial Fibrillation") %>% 
  clean_names() %>% 
  filter(ind == "AF07") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2012-13", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)

# Repeat for 2011/12 indicator "AF03"
# **Note change in definition here - includes anti-platelets**
af_1112 <-
  read_excel(here("03_data/QOF_2011_12/Individual level 2011-12/QOF_Scot_201112_indiv_af.xls"),
           sheet = "Atrial Fibrillation") %>%
  clean_names() %>% 
  filter(ind == "AF03") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2011-12", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)

# Repeat for 2010/11 indicator "AF03"
# **Note change in definition here - includes anti-platelets**
# **Note folder structure says 2009/10 but actual files are for 2010/11**

af_1011 <-
  read_excel(here("03_data/QOF_2009_10/Individual level 2009-10/QOF_Scot_201011_indiv_af.xls"),
           sheet = "Atrial Fibrillation") %>% 
  clean_names() %>% 
  filter(ind == "AF03") %>% 
  select(practice, indicator = ind, ind_num = num, ind_den = denom) %>% 
  mutate(fin_year = "2010-11", indicator = "af007", 
         practice = as.character(practice),
         across(ind_num:ind_den, as.double), ind_perc = ind_num/ind_den * 100) %>% 
  round_df(dig = 1) %>% 
  left_join(., pop_table)
```

Now combine all 

```{r}
# Combine all AF data together
af_tab <- 
  full_join(tqa_ind, af_1516) %>% 
  full_join(af_1415) %>% 
  full_join(af_1314) %>% 
  full_join(af_1213) %>% 
  full_join(af_1112) %>% 
  full_join(af_1011) %>% 
  arrange(indicator, practice, fin_year) %>% 
  group_by(practice) %>% 
  fill(name, .direction = "up") %>% 
  fill(board, .direction = "up") %>% 
  ungroup

af_tab
```



#### Registers

Think about how to do this - register info already inputted then dropped - work around that.....

### BP

### CHD

### CKD

### DM

### HF

### HYP

### MH

### PAD

###STIA

## Combine


## SIMD and UR data






